---
title: 'Project #1'
author: "Emily Byrne"
date: "2022-10-06"
output: html_document
---

## Project 1: Analysis of Customers Who Buy Diet Products and/or Candy {.tabset}

### Packages

```{r}
library(ggplot2) #gives us the ability to plot our data
library(completejourney) #where our data is located
library(dplyr) #allows us to manipulate our data
library(stringr) #allows us to work with character strings
library(lubridate) #allows us to work with dates and times
library(tidyverse) #allows us to clean up our data, makes it easier to read
transactions <- get_transactions()
```
### Tables
```{r}
transactions
products
demographics
```

### Tables for Groups
  We created three separate data tables for those households who purchase both candy and diet products, households who purchases either candy or diet products, and households who purchase neither. 

```{r}
both <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  filter(str_detect(product_type, regex("DIET", ignore_case = TRUE)),
         str_detect(product_category, regex("CANDY", ignore_case = TRUE)))

either <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  filter(str_detect(product_type, regex("DIET", ignore_case = TRUE))|
         str_detect(product_category, regex("CANDY", ignore_case = TRUE)))

neither <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  filter(!str_detect(product_type, regex("DIET", ignore_case = TRUE)),
         !str_detect(product_category, regex("CANDY", ignore_case = TRUE)))

```

```{r}
both
either
neither
```

### Distinct Households and their Purchasing Behavior Table
  Here we mutated the original data table to contain a new variable called "Diet_vs._Candy" to tell which group the household belongs in. We also cut the data frame down to show only the distinct households IDs and the groups. 
```{r}
distinct_households <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy) %>% 
  summarize(distinct_household_id = unique(household_id))

distinct_households
```
### Filtered Tables for Count of Groups
  We filtered the table to show only the results for each group seperateluy to give us an accurate count of the amount of distinct households in each group. 
```{r}
distinct_households %>% 
  filter(Diet_vs._Candy == "Both") %>%
  count()

distinct_households %>% 
  filter(Diet_vs._Candy == "Either") %>%
  count()

count_neither <- distinct_households %>% 
  filter(Diet_vs._Candy == "Neither") %>%
  count()
```

### Graph grouped by age for Both Diet & Candy
```{r}
group_by_both <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy, age) %>% 
  filter(Diet_vs._Candy == "Both") %>%
  summarize(distinct_household_id = unique(household_id)) %>%
  ggplot(aes(x = age, fill = age)) +
  geom_bar() +
  ggtitle("The Age Range That Bought Both Diet Products & Candy Together", subtitle = "The data below shows which age range has the tendancy 
 to buy both Diet products and Candy together throughout all transactions") +
  xlab("Age Range") +
  scale_y_continuous(name = "Number of Transactions") +
  scale_fill_discrete("Age Range") +
  coord_flip()

group_by_both +
  theme(plot.title = element_text(size = 13, face = "bold")) +
  theme(legend.position = "none") +
  theme(plot.subtitle = element_text(color = "azure4", size = 8)) +
  theme(panel.background = element_rect(fill = "oldlace"))
  
```

### Graph grouped by age for Either Diet or Candy
```{r}
group_by_either <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy, age) %>% 
  filter(Diet_vs._Candy == "Either") %>%
  summarize(distinct_household_id = unique(household_id)) %>%
  ggplot(aes(x = age, fill = age)) +
  geom_bar() +
  ggtitle("The Age Range That Bought Either Diet Products or Candy", subtitle = 
"The data below shows which age range has the tendancy 
 to buy either Diet products or Candy but not together throughout all transactions") +
  xlab("Age Range") +
  scale_y_continuous(name = "Number of Transactions") +
  scale_fill_discrete("Age Range") +
  coord_flip()

group_by_either +
  theme(plot.title = element_text(size = 13, face = "bold")) +
  theme(legend.position = "none") +
  theme(plot.subtitle = element_text(color = "azure4", size = 8)) +
  theme(panel.background = element_rect(fill = "oldlace"))
```

### Graph grouped by age for Neither Diet nor Candy
```{r}
group_by_neither <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy, age) %>% 
  filter(Diet_vs._Candy == "Neither") %>%
  summarize(distinct_household_id = unique(household_id)) %>%
  ggplot(aes(x = age, fill = age)) +
  geom_bar() +
  ggtitle("The Age Range That Bought Neither Diet Products nor Candy", subtitle = "The data below shows which age range has the tendancy 
 to buy neither Diet products nor Candy throughout all transactions") +
  xlab("Age Range") +
  scale_y_continuous(name = "Number of Transactions") +
  scale_fill_discrete("Age Range") +
  coord_flip()

group_by_neither +
  theme(plot.title = element_text(size = 13, face = "bold")) +
  theme(legend.position = "none") +
  theme(plot.subtitle = element_text(color = "azure4", size = 8)) +
  theme(panel.background = element_rect(fill = "oldlace"))
```


### Graph grouped by income for Both Diet and Candy
```{r}
income_by_both <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy, income) %>% 
  filter(Diet_vs._Candy == "Both") %>%
  summarize(distinct_household_id = unique(household_id)) %>%
   ggplot(aes(x = income, fill = income)) +
  geom_bar() +
  ggtitle("The Income Range That Bought Both Diet Products & Candy Together", subtitle = "The data below shows which income range has the tendancy 
 to buy both Diet products and Candy throughout all transactions") +
  xlab("Income Range") +
  scale_y_continuous(name = "Number of Transactions") +
  scale_fill_discrete("Income Range") +
  coord_flip()

income_by_both +
  theme(plot.title = element_text(size = 13, face = "bold")) +
  theme(legend.position = "none") +
  theme(plot.subtitle = element_text(color = "azure4", size = 8)) +
  theme(panel.background = element_rect(fill = "oldlace"))
  
```

### Graph grouped by income for Either Diet or Candy
```{r}
income_by_either <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy, income) %>% 
  filter(Diet_vs._Candy == "Either") %>%
  summarize(distinct_household_id = unique(household_id)) %>%
   ggplot(aes(x = income, fill = income)) +
  geom_bar() +
  ggtitle("The Income Range That Bought Either Diet Products or Candy", subtitle = "The data below shows which income range has the tendancy 
 to buy either Diet products or Candy throughout all transactions") +
  xlab("Income Range") +
  scale_y_continuous(name = "Number of Transactions") +
  scale_fill_discrete("Income Range") +
  coord_flip()

income_by_either +
  theme(plot.title = element_text(size = 13, face = "bold")) +
  theme(legend.position = "none") +
  theme(plot.subtitle = element_text(color = "azure4", size = 8)) +
  theme(panel.background = element_rect(fill = "oldlace"))
```

### Graph grouped by income for Neither Diet nor Candy
```{r}
income_by_neither <- products %>% 
  inner_join(transactions) %>% 
  inner_join(demographics) %>% 
  mutate(Diet_vs._Candy = ifelse(grepl("DIET", product_type) & grepl("CANDY", product_category), "Both", ifelse(grepl("DIET", product_type) | grepl("CANDY", product_category), "Either", "Neither"))) %>% 
  group_by(Diet_vs._Candy, income) %>% 
  filter(Diet_vs._Candy == "Neither") %>%
  summarize(distinct_household_id = unique(household_id)) %>%
   ggplot(aes(x = income, fill = income)) +
  geom_bar() +
  ggtitle("The Income Range That Bought Neither Diet Products nor Candy", subtitle = "The data below shows which income range has the tendancy 
 to buy neither Diet products nor Candy throughout all transactions") +
  xlab("Income Range") +
  scale_y_continuous(name = "Number of Transactions") +
  scale_fill_discrete("Income Range") +
  coord_flip()

income_by_neither +
  theme(plot.title = element_text(size = 13, face = "bold")) +
  theme(legend.position = "none") +
  theme(plot.subtitle = element_text(color = "azure4", size = 8)) +
  theme(panel.background = element_rect(fill = "oldlace"))
```












